### 2015-03-15 Paper.tex
Possibly final edited version.

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}
\usepackage{multirow}
\graphicspath{ {images/} }

\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{
	language=Java,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numbers=none,
	numberstyle=\tiny\color{gray},
	keywordstyle=\color{blue},
	commentstyle=\color{dkgreen},
	stringstyle=\color{mauve},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3
}

\begin{document}
\title{Hadoop Image Processing Framework}
\author{\IEEEauthorblockN{Sridhar Vemula}
  \IEEEauthorblockA{Computer Science Department\\
    Oklahoma State University\\
    Stillwater, Oklahoma \\
    Email: sridhar.vemula@okstate.edu}
  \and
  \IEEEauthorblockN{Christopher Crick}
  \IEEEauthorblockA{Computer Science Department\\
    Oklahoma State University\\
    Stillwater, Oklahoma \\
    Email: chriscrick@cs.okstate.edu}
}
\maketitle
\begin{abstract}
With the rapid growth of social media, the number of images being
uploaded to the internet is exploding. Massive quantities of images
are shared through multi-platform services such as Snapchat,
Instagram, Facebook and WhatsApp; recent studies estimate that over
1.8 billion photos are uploaded every day.  However, for the most
part, applications that make use of this vast data have yet to
emerge. Most current image processing applications, designed for
small-scale, local computation, do not scale well to web-sized
problems with their large requirements for computational resources and
storage.  The emergence of processing frameworks such as the Hadoop
MapReduce\cite{dean2008} platform addresses the problem of providing
a system for computationally intensive data processing and distributed
storage. However, to learn the technical complexities of developing
useful applications using Hadoop requires a large investment of time
and experience on the part of the developer.  As such, the pool of
researchers and programmers with the varied skills to develop
applications that can use large sets of images has been limited. To
address this we have developed the Hadoop Image Processing Framework,
which provides a Hadoop-based library to support large-scale image
processing. The main aim of the framework is to allow developers of
image processing applications to leverage the Hadoop MapReduce
framework without having to master its technical details and introduce
an additional source of complexity and error into their programs.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
With the spread of social media in recent years, a large amount of
image data has been accumulating. When processing this massive data
resource has been limited to single computers, computational power and
storage ability quickly become bottlenecks. Alternately, processing
tasks can typically be performed on a distributed system by dividing
the task into several subtasks. The ability to parallelize tasks
allows for scalable, efficient execution of resource-intensive
applications.  The Hadoop MapReduce framework provides a platform for
such tasks.

When considering operations such as face detection, image
classification\cite{Li2009} and other types of processing on images,
there are limits on what can be done to improve performance of single
computers to make them able to process information at the scale of
social media. Therefore, the advantages of parallel distributed
processing of a large image dataset by using the computational
resources of a cloud computing environment should be considered. In
addition, if computational resources can be secured easily and
relatively inexpensively, then cloud computing is suitable for
handling large image data sets at very low cost and increased
performance. Hadoop, as a system for processing large numbers of
images by parallel and distributed computing, seems promising. In
fact, Hadoop is in use all over the world. Studies using Hadoop have
been performed, dealing with text data files\cite{Lin2010}, analyzing
large volumes of DNA sequence data\cite{McKenna2010}, converting the
data of a large number of still images to PDF format, and carrying out
feature selection/extraction in astronomy\cite{wiley2011}.  These
examples demonstrate the usefulness of the Hadoop system, which can
run multiple processes in parallel for load balancing and task
management.

Most of the image processing applications that use the Hadoop
MapReduce framework are highly complex and impose a staggering
learning curve.  The overhead, in programmer time and expertise,
required to implement such applications is cumbersome.  To address
this, we present the Hadoop Image Processing Framework, which hides
the highly technical details of the Hadoop system and allows
programmers who can implement image processing algorithms but who have
no particular expertise in distributed systems to nevertheless
leverage the advanced resources of a distributed, cloud-oriented
Hadoop system. Our framework provides users with easy access to
large-scale image data, smoothly enabling rapid prototyping and
flexible application of complex image processing algorithms to very
large, distributed image databases.

The Hadoop Image Processing Framework is largely a software
engineering platform, with the goal of hiding Hadoop's complexity
while providing users with the ability to use the system for
large-scale image processing without becoming crack Hadoop engineers.
The framework's ease of use and Java-oriented semantics will further
ease the process of creating large scale image applications and
experiments. This framework is an excellent tool for novice Hadoop
users, image application developers and computer vision researchers,
allowing the rapid development of image software that can take
advantage of the huge data stores, rich metadata and global reach of
current online image sources.\footnote[1]{The Hadoop Image Processing
  Framework is open source and freely available for download at
  https://github.com/okstate-robotics/hipl.git}.

In the following section we will describe prior work in this
area. Next we present an overview of the Hadoop Image Processing
Framework including the Downloader, Processor and Extractor
stages. Additionally, we describe our approach of distributing tasks
for MapReduce. Finally, we demonstrate the potential of this framework
with quantitative analysis and experiments performed on image
processing tasks.

\section{Related Work}
With the rapid usage increase of online photo storage and social media
on sites like Facebook, Flickr and Picasa, more image data is
available than ever before, and is growing every day.  Every minute
27,800 photos are uploaded to Instagram,\cite{Horaczek2013} while
Facebook receives 208,300 photos over the same time frame. This alone
provides a source of image data that can scale into the billions.  The
explosion of available images on social media has motivated image
processing research and application development that can take
advantage of very large image data stores.

White et.al \cite{White2010} presents a case study of classifying and
clustering billions of regular images using MapReduce.  It describes
an image pre-processing technique for use in a sliding-window approach
for object recognition.  Pereira et.al \cite{Pereira2010} outlines
some of the limitations of the MapReduce model when dealing with
high-speed video encoding, namely its dependence on the NameNode as a
single point of failure, and the difficulties inherent in generalizing
the framework to suit particular issues.  It proposes an alternate
optimized implementation for providing cloud-based IaaS
(Infrastructure as a Service) solutions.  Lv et.al \cite{Lv2010}
describes using the $k$-means algorithm in conjunction with MapReduce
and satellite/aerial photographs in order to find different elements
based on their color.

Zhang et.al \cite{Zhang2010} presents methods used for processing
sequences of microscope images of live cells. The images are
relatively small (512x512, 16-bit pixels).  Stored in 90 MB folders,
the authors encountered difficulties regarding fitting into Hadoop DFS
blocks with were solved by custom InputFormat, InputSplit and
RecordReader classes.  Powell et.al \cite{Powell2010} describes how
NASA handles image processing of celestial images captured by the Mars
Orbiter and rovers. Clear and concise descriptions are provided about
the segmentation of gigapixel images into tiles, how the tiles are
processed and how the image processing framework handles scaling and
works with the distributed processing. Wang, Yinhai and
McCleary\cite{Wang2011} discuss speeding up the analysis of tissue
microarray images by substituting human expert analysis for automated
processing algorithms. While the images were gigapixel-sized, the
content was easily segmented and there was no need to analyze all of
an image at once. The work was all done on a specially-built high
performance computing platform using the Hadoop framework.

Bajcsy et.al \cite{Bajcsy2013} present a characterization of four
basic terabyte-size image computations on a Hadoop cluster in terms of
their relative efficiency according to a modified Amdahl's Law. The
work was motivated by the fact that there is a lack of standard
benchmarks and stress tests for large-scale image processing
operations on the Hadoop framework. Moise et.al \cite{Moise2013}
outlines the querying of thousands of images in one run using the
Hadoop MapReduce framework and the eCP Algorithm. The experiment
performs an image search on 110 million images collected from the web
using the Grid 5000 platform. The results are evaluated in order to
understand the best practices for tuning Hadoop MapReduce performance
for image search.

The above shows that there has been a great deal of intensive work in
image processing using MapReduce.  However, each independent project
requires a complex, error-prone, one-off implementation.  Such
research and application development would benefit greatly from a
standard, well-engineered image processing framework such as the one
we provide.

\section{Methodology}
\label{methodology}
The Hadoop Image Processing Framework is intended to provide users
with an accessible, easy-to-use tool for developing large-scale image
processing applications.

The main goals of the Hadoop Image Processing Framework are:
\begin{itemize}
\item Provide an open source framework over Hadoop MapReduce for
  developing large-scale image applications
\item Provide the ability to flexibly store images in various Hadoop
  file formats
\item Present users with an intuitive application programming
  interface for image-based operations which is highly parallelized
  and balanced, but which hides the technical details of Hadoop
  MapReduce
\item Allow interoperability between various image processing
  libraries
\end{itemize}

\subsection{Downloading and storing image data}
Hadoop uses the Hadoop Distributed File System (HDFS)\cite{Shvachko2010} to store files
in various nodes throughout the cluster.  One of Hadoop's significant
problems is that of small file storage. \cite{White2009} A small file
is one which is significantly smaller than HDFS block size.  Large
image datasets are made up of small image files in great numbers,
which is a situation which HDFS has a great deal of trouble
handling. The problem can be solved by providing a container to group
the files in some way. Hadoop offers a few options:
\begin{itemize}
	\item HAR File
	\item Sequence File
	\item Map File
\end{itemize}

The Downloader Module of our Hadoop Image Processing Framework
performs the following operations:

\textbf{Step 1: Input a URL List.} Initially users input a file
containing URLs of images to download. The input list should be a text
file with one image URL per line. The list can be generated by hand,
extracted from a database or a provided by a search. The framework
provides a extendable ripper module for extracting URLS from Flickr
and Google image searches and from SQL databases. In addition to the
list the user selects the type of image bundle to be generated
(e.g. HAR, sequence or map).  Our system divides the URLs for download
across the available processing nodes for maximum efficiency and
parallelism.  Each node map task generates several image bundles
appropriate to the selected input type, containing all of the images
it downloaded. In the reduce phase, the Reducer will merge these image
bundles into a large image bundle.

\textbf{Step 2: Split URLs across nodes.}  From the input file
containing the list of image URLs and the type of file to be
generated, we equally distribute the task of downloading images across
the all the nodes in the cluster. The nodes are efficiently managed so
that no memory overflow can occur even for terabytes of images
downloaded in a single map task. This allows maximum downloading
parallelization. Image URLs are distributed among all available
processing nodes, and each map task begins downloading its respective
image set.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{down-map}
	\caption{Individual map task of Downloader Module}
	\label{fig:down-map}
\end{figure}

\textbf{Step 3: Download image data from URLs.}  For every URL
retrieved in the map task, a connection is established according to
the appropriate transfer protocol (e.g. FTP, HTTP, HTTPS, etc.). Once
connected, we check the file type.  Valid images are assigned
InputStreams associated with the connection. From these InputStreams,
we generate new HImage objects and add the images to the image
bundle. The HImage class holds the image data and provides an
interface for the user's manipulation of image and image header
data. The HImage class also provides interoperability between various
image data types (e.g. BufferedImage, Mat, etc.).

\textbf{ Step 4: Store images in an image bundle. }  Once an HImage
object is received, it can be added to the image bundle simply by
passing the HImage object to the appendImage method. Each map task
generates a number of image bundles depending on the image list. In
the reduce phase, all of these image bundles are merged into one large
bundle.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{down-node}
	\caption{Single node running the Downloader Module (handled by
          the framework and transparent to the user)}
	\label{fig:down-node}
\end{figure}

\subsection{Processing image bundle using MapReduce}
Hadoop MapReduce programs handle input and output data very
efficiently, but their native data exchange formats are not convenient
for representing or manipulating image data.  For instance,
distributing images across map nodes require the translation of images
into strings, then later decoding these image strings into specified
formats in order to access pixel information.  This is both
inefficient and inconvenient. To overcome this problem, images should
be represented in as many different formats as possible, increasing
flexibility. The framework focuses on bringing familiar data types
directly to user.

As distribution is important in MapReduce, images should be processed
in the same machine where the bundle block resides. In a generic
MapReduce system, the user is responsible for creating InputFormat and
RecordReader classes to specify the MapReduce job and distribute the
input among nodes. This is a cumbersome and error-prone task; the
Hadoop Image Processing Framework provides such InputFormat and
RecordReaders for the sake of the system's ease of use.

Images are distributed as various image data types and users have
immediate access to pixel values.  If pixel values are naively
extracted from the image byte formats, valuable image header data
(e.g. JPEG EXIF data or IHDR \cite{David03} image headers) is
lost. The framework holds the image data in a special HImageHeader
data type before converting the image bytes into pixel values.  After
processing pixel data, image headers are reattached to the processed
results.  The small amount of storage overhead required for this
functionality is a trade-off worth making for the sake of preserving
image header data after processing.

The functionality of the framework's Processor module is described
below:

\textbf{Step 1: Devise the algorithm.} We assume that the user writes
an algorithm which extends the provided GenericAlgorithm class. This
class is passed as an argument to the processor module. The framework
starts a MapReduce job with the algorithm as an input. The
GenericAlgorithm holds an HImage variable; this allows user to write
an algorithm on a single image data structure, which the framework
then iterates over the entire image bundle. In addition to the
algorithm, the user should provide the image bundle file that needs to
be processed.  Depending on the specifics of the image bundle
organization and contents, the bundle is divided across nodes as
individual map tasks. Each map task will apply the processing
algorithm to each local image and append them in turnto the output
image bundle. In the reduce phase, the Reducer merges these image
bundles into a large image bundle.

\textbf{Step 2: Split image bundle across nodes.} The input image
bundle is stored as blocks in the HDFS.  In order to obtain maximum
throughput, the framework establishes each map task to run in the same
block where it resides, using custom input format and record reader
classes. This allows maximum parallelization without the problem of
transferring data across nodes.  Each image bundle now applies
different map tasks to the image data for which it is responsible.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{pro-map}
  \caption{Individual map task of Processor Module}
  \label{fig:pro-map}
\end{figure}

\textbf{Step 3: Process individual image.}  The processing algorithm
devised by the user and provided as input to the Processing Module is
applied to every HImage retrieved in the map task.  The HImage
provides its image data in the data format (e.g. Java BufferedImage,
OpenCV Mat, etc.) requested by the user and used by the processing
algorithm. Once the image data type is retrieved, processing takes
place. After processing, the preserved image header data from the
original image is appended to the processed image. The processed image
is appended to the temporary bundle generated by the map task.

\textbf{Step 4: Store processed images in an image bundle.} Every map
task generates an image bundle upon completion of its processing.
Once the map phase is completed there are many bundles scattered
across the computing cluster.  In the reduce phase, all of these
temporary image bundles are merged into a single large file which
contains all the processed images.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{pro-node}
  \caption{Single node running the Processor Module (handled by
    the framework and transparent to user)}
  \label{fig:pro-node}
\end{figure}

\subsection{Extracting image bundles using MapReduce}
In addition to creating and processing image bundles, the framework
provides a method for extracting and viewing these images.  Generally,
Hadoop extracts images from an image bundle iteratively, inefficiently
using a single node for the task. To address this inefficiency, we
designed an Extractor module which extracts images in parallel across
all available nodes.

Distribution plays a key role in MapReduce; we want to make effective
and efficient use of the nodes in the computing cluster.  As
previously mentioned in the description of the Processor module, a
user working in a generic Hadoop system must again devise custom
InputFormat and RecordReader classes in order to facilitate
distributed image extraction.  The Hadoop Image Processing Framework
provides this functionality for the extraction task as well, providing
much greater ease of use for the development of image processing
applications.

Organizing and specifying the final location of extracted images in a
large distributed task can be confusing and difficult.  Our framework
provides this functionality, and allows the user simply to specify
whether images should be extracted to a local file system or reside on
the Hadoop DFS.

The process of the Extractor module is explained below:

\textbf{Step 1: Input the image bundle to be extracted.} The image
bundle specified for extraction is passed as a parameter to the
Extractor module.  In addition, the user can include an optional
parameter specifying the images' final location (defaults to local
file system).  The image bundle is then distributed across the nodes
as individual map tasks. Each map task will extract the requisite
images onto the specified filesystem.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{ext-map}
  \caption{Individual map task of Extractor Module}
  \label{fig:ext-map}
\end{figure}

\textbf{Step 2: Split Image bundle across nodes} The input image
bundle is split across available nodes using the framework's custom
input format and record reader classes for maximum throughput. Once a
map task starts, HImage objects are retrieved.

\textbf{Step 3: Extract individual image} The image bytes obtained
from each HImage object are stored onto the filesystem in the
appropriate file format based on its image type.  The Extractor module
lacks a reduce phase.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{ext-node}
  \caption{Single node running the Extractor Module (handled by the
    framework and transparent to the user)}
  \label{fig:ext-node}
\end{figure}

\section{Image Processing Algorithms}
\label{algorithms}
To explore the Hadoop Image Processing Framework's capabilities and
performance, we implemented multiple variations of existing
widely-used image processing algorithms:
\begin{itemize}
\item A stand-alone implementation running on a single node with no
  distributed processing capability
\item A generic Hadoop implementation which requires a high level of
  software engineering expertise within the Hadoop framework
\item An implementation that employs the Hadoop Image Processing
  Framework
\end{itemize}

We chose the following algorithms: Laplacian filtering, Canny edge
detection and $k$-means image segmentation.  These are widely-used,
computation- and data-intensive algorithms of varying complexity which
require large distributed systems for timely operation on hundreds of
thousands of images.

%\subsection{Flat Field Correction}
%Flat Field Correction(FFC) is described mathematically below:
%\begin{equation}
%	I^{FFC}(x,y) = \frac{I^{RAW}(x,y)-DI(x,y)}{WI(x,y)-DI(x,y)}
%\end{equation}

%where \begin{math}I^{FFC}(x,y))\end{math} is the flat-field corrected image intensity, \begin{math}DI(x,y)\end{math} is the dark image acquired by closing the camera shutter, \begin{math}I^{RAW}(x,y)\end{math} is the raw uncorrected image density and WI(x,y) is the flat field intensity acquired without any object to correct primarily for spatial shading. This is the simpleset computations that consists of two subtractions and one division for pixel. It needs the DI and WI images co-located from the distributed execution perspective.

\subsection{Laplacian Filter}
The Laplacian is a 2-D isotropic measure of the 2nd spatial derivative
of an image. The Laplacian of an image highlights regions of rapid
intensity change and is therefore often used for edge detection. The
Laplacian is often applied to an image that has first been smoothed
with an approximation of a Gaussian filter in order to reduce
sensitivity to noise. The operator normally takes a single gray level
image as input and produces another gray level image as output.

The Laplacian \begin{math} L(x,y)\end{math}of an image with pixel
intensity values \begin{math}I(x,y)\end{math} is given by
\begin{equation}
L(x,y) = \frac{\partial^2 I}{\partial x^2} + \frac{\partial^2 I}{\partial y^2}
\end{equation}

This is calculated using a convolution filter.

Since the input image is represented as a set of discrete pixels, we
have to find a discrete convolution kernel that can approximate the
second derivatives in the definition of the Laplacian. This is a very
simple computation that consists of a few additions and
multiplications.

\subsection{Canny Edge Detection}
Canny edge detection\cite{Canny86} is an multi-stage algorithm to
detect a wide range of edges in images. Edge detection, especially
step edge detection, is an important technique to extract useful
structural information from visual information, dramatically reducing
the amount of data to be processed. Among existing edge detection
methods, the Canny edge detection algorithm is considered to provide
reliable edge detection without relying on context-specific
heuristics.

The process of the Canny edge detection algorithm can defined in five
stages:

\begin{itemize}
\item Apply a Gaussian filter to smooth the image and remove noise.
\item Find the intensity gradients of the image.
\item Apply non-maximum suppression to get rid of spurious
  responses to edge detection
\item Apply a double threshold to determine potential edges.
\item Track edges by hysteresis. Finalize the detection of
  edges by suppressing all edges that are ``weak'' and not
  connected to strong edges.
\end{itemize}

\subsection{Image segmentation using $k$-means clustering}
The $k$-means algorithm is an unsupervised clustering algorithm that
classifies input data points into multiple classes based on their
Euclidean distance from each other. The algorithm assumes that the
data features form a vector space and tries to find natural
clusterings within it. The points are clustered around
centroids \begin{math}\mu_{i} \forall i = 1...k\end{math} which are
  obtained by minimizing the objective
\begin{equation}
	V = \sum\limits_{i=1}^n \sum\limits_{x_{j}\in S_{i}} (x_{j} - \mu_{i})^2
\end{equation}

where there are \begin{math}k\end{math} clusters \begin{math}
    S_{i},i=1,2,...,k \end{math} and \begin{math} \mu_{i} \end{math}
  is the centroid or mean point of all the points \begin{math}x_{j}
    \in S_{i}\end{math}.

For the purposes of our experiment, we implemented iterative versions
of the algorithm. The algorithm takes a 2-dimensional image as
input. The steps of the algorithm are as follows:

\begin{itemize}
\item Compute the intensity distribution (also called the
  histogram) of the image.
\item Initialize the centroids with $k$ random intensities.
\item Repeat the following steps until the cluster labels of
  the image stop changing from one iteration to the next:
  \begin{itemize}
  \item Cluster the points based on the distance of their intensities
    from the centroid's intensity.
	\begin{equation}
		c^{i} := \mbox{argmin}_{j}||x^{i} - \mu_{j}||^2
	\end{equation}
	\item Compute the new centroid for each of the clusters.
	\begin{equation}
		\mu_{i} = \frac{\sum\limits_{i=1}^{m} 1\{c_{i} = j\}x^i}{\sum\limits_{i=1}^{m} 1\{c_{i} = j\}}
	\end{equation}

	where \begin{math}k\end{math} is a parameter of the algorithm
          (the number of clusters to be
          found), \begin{math}i\end{math} iterates over all the
            intensities, \begin{math}j\end{math} iterates over all the
              centroids and \begin{math}\mu_{i}\end{math} are the
              centroid intensities.
  \end{itemize}
\end{itemize}

\section{Experimental Results}

We present execution time and programming complexity results for the
three image processing algorithms described above, demonstrating the
effectiveness of the Hadoop Image Processing Framework.  These
algorithms are not particularly challenging to implement on a single
node, but producing a distributed Hadoop implementation is quite
challenging and requires a great deal of familiarity with Hadoop
concepts and distributed software engineering expertise.  On the other
hand, our framework provides all of the performance and scalability
benefits of the Hadoop system while retaining the simplicity of the
single-node implementation.  In this section we describe the dataset
and hardware and software specifications used in our experiments, and
discuss the results.

%\begin{figure}[h]
%	\centering
%	\includegraphics[width=0.45\textwidth]{input-canny}
%	\caption{A Sample Image from downloaded image bundle}
%	\label{fig:input-canny}
%\end{figure}

%\begin{figure}[h]
%	\centering
%	\includegraphics[width=0.45\textwidth]{output-canny}
%	\caption{Processed image using canny edge detection}
%	\label{fig:output-canny}
%\end{figure}

\subsection{Characteristics of Image Dataset, Hardware and Software}
We performed image processing computations on a dataset of about 1TB
in size. The dataset was obtained by performing a Flickr search based
on the keyword ``flight''.  The dataset was extracted using the
FlickrRipper, an instance of the Ripper interface provided by our
Hadoop Image Processing Framework.  The framework currently provides
rippers for Flickr and Google Image searches, and more will be
provided in the future.  A Ripper interface extracts image URLs from
search results, and can be implemented for any set of results in any
format.  The experimental dataset is composed of 220000 images at 4.76
MB/image, or about 1 TB.

\subsection{Computer Hardware and Software Characterstics}
We ran the image processing computations on a 6-node distributed
cluster and on a single-node desktop computer. The table below
summarizes the cluster and desktop hardware and software
specifications. The cluster nodes differ in terms of CPU speed and
RAM. We installed Hadoop, Cloudera CDH5 and Java 1.7 on the cluster to
support Java code execution. The desktop computer had a similar
software configuration to the cluster.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 data nodes with \\ from 2 to 4 virtual \\ processors and 4 GB\\ RAM and name node \\ with 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyperthreading\\ activated\end{tabular} \\ \cline{2-4} 
		& Networking                                                     & 1 Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4} 
		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}Hadoop 2.5\\ Cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4} 
		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
	\end{tabular}
\end{table}

The desktop implementations of the algorithms are Java programs for
processing a small image dataset. As soon as the data set is processed
the program halts. The desktop implementation starts as a single
thread, and images are processed one at a time in succession.  When
the dataset size is low, each program runs smoothly and with excellent
performance.  However, as the size of the data to be processed reaches
the terabyte scale, such a single-process approach fails utterly.
Data on this scale must be processed in distributed fashion using
Hadoop.

Each algorithm is also implemented on a Hadoop cluster without using
the Hadoop Image Processing Framework.  These implementations overcome
the problem of handling large datasets but they suffer from extremely
complex code.  The user needs to write a great deal of code and
understand the technical details of Hadoop. Writing such code,
including custom InputFormats and RecordReaders, is cumbersome and
error-prone. Fig. \ref{fig:module-chart} compares the actual lines of
code a user needs to write in order to run our experimental algorithms
(or any similar image processing application) on a Hadoop cluster.

Finally, we compare the same algorithms implemented on a Hadoop
cluster using the framework.  These implementations overcome the
problems of handling large datasets and managing interoperability
between image datatypes while holding code complexity down to the same
level as the single-process, local implementation.  Users enjoy
automated creation of custom InputFormats and RecordReaders, and can
tune settings for the desired output.  The framework is so transparent
that the user needs only a few lines of code to process a large image
data set. It hides the technical details of MapReduce and runs the
code in highly parallelized fashion.  The code required to implement
each image processing algorithm is no more complex than the same
algorithm implemented on a local, single-threaded system, but the
performance improvement for large datasets is immense.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		\multicolumn{1}{|c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computa-\\ tional\\ Platform\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computational \\ Elasticity\end{tabular}}}      & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Code \\ Complexity\end{tabular}}}                                                                                                                                                  & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Data \\ Locality\end{tabular}}}                                                                           \\ \hline
		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies in writing every \\ module in framework \\ including custom \\ InputFormats and \\ RecordReaders.  User \\ requires high \\ technical expertise \\ with Hadoop.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS. Custom\\ InputFormats \\ need to be \\ created to launch\\ computations \\ in appropriate \\ locations\end{tabular} \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ on desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS. Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
	\end{tabular}
\end{table}

\iffalse
\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		\textbf{Module} & \textbf{\begin{tabular}[c]{@{}c@{}}SLOC \\ using framework\end{tabular}}                                                          & \textbf{\begin{tabular}[c]{@{}c@{}}SLOC \\ without framework\end{tabular}}                                                                                    \\ \hline
		Downloader      & 3 - 5 lines                                                                                                                       & \begin{tabular}[c]{@{}c@{}}300 - 1000 lines.\\ Custom Input Formats\\ and RecordReaders are\\ to be implemented\end{tabular}                                  \\ \hline
		Processor       & 3 - 5 lines                                                                                                                       & \begin{tabular}[c]{@{}c@{}}400 - 1000 lines\\ Many overheads. \\ Converting to image \\ data types, Mapper \\ locality should be taken\\ care of\end{tabular} \\ \hline
		Extractor       & 2 - 8 lines                                                                                                                       & \begin{tabular}[c]{@{}c@{}}200 - 1000 lines\\ Depends on the extraction\\ process.\end{tabular}                                                               \\ \hline
		Algorithm       & \begin{tabular}[c]{@{}c@{}}200 - 300 lines\\ adds a few lines overhead\\ to automate the process\\ for MapReduce job\end{tabular} & 200 - 250 lines                                                                                                                                               \\ \hline
	\end{tabular}
\end{table}
\fi

\subsection{Observations}

We have explored the space of image processing algorithms, applying
different hardware and software configurations to the various data
sets.  Specifically, we compared:

\begin{itemize}
\item Three image processing algorithms described in Section
  \ref{algorithms}
\item Datasets of varying image counts
\item Configurations of hardware cluster used for computations 
\item Coding complexity (measured in Lines of Code (LOC)
\end{itemize}

The numerical results are shown in Figures \ref{fig:perf-chart}
through \ref{fig:files-chart}.

Figure \ref{fig:perf-chart} illustrates the performance comparison of
Canny edge detection on a single node, on Hadoop using our framework
and Hadoop without using the framework.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{perf-graph}
  \caption{Canny edge detection computation executed on a single node
    and on a Hadoop cluster with and without using framework}
  \label{fig:perf-chart}
\end{figure}

It can be observed that the Hadoop cluster performance is nearly
identical whether the Hadoop Image Processing Framework is used or
not.  The single node performance is satisfactory on small input
sizes; in fact it outperforms Hadoop when the amount of data is very
small owing to the lack of MapReduce overhead.  However, as the amount
of data grows the single node performance becomes increasingly poor.
The same pattern is observed with any image processing algorithm.

Figure \ref{fig:module-chart} compares the coding complexity of the
different image handling and processing tasks described in Section
\ref{methodology}, for the Canny edge detection task. The Downloader
module and Extractor module are simple to use and require few lines of
code to configure. The processor module includes the specific
algorithm devised by the user, which can be significant amount of
code, but using our framework to configure the algorithm for MapReduce
requires almost no additional coding.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{module-chart2}
  \caption{Comparing coding complexities of modules in different
    environments}
  \label{fig:module-chart}
\end{figure}

Figure \ref{fig:comp-chart} shows the coding complexity of the
processor portion of our three experimental image processing
algorithms.  The naive single-process, single-node algorithm requires
a similar amount of code to using our highly parallel, distributed
framework, while direct implementation of the algorithm using Hadoop
and MapReduce requires a great deal more effort.  The Hadoop Image
Processing Framework hides all of this complexity from the user.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{comp-chart}
  \caption{Comparing coding complexity of Laplacian filter,
    Canny edge detection and $k$-means clustering in different
    environments}
  \label{fig:comp-chart}
\end{figure}

The Hadoop Image processing framework provides transparency on
multiple levels.  Users can use the predefined MapReduce modules
(Downloader, Processor and Extractor) for processing the image
modules, and most image processing tasks require no more than this.
Users may also write custom MapReduce tasks within the framework,
taking advantage of its hierarchical construction.  This still
protects the user from the full complexity and error-prone nature of
the full Hadoop framework, but requires additional knowledge and
expertise on the user's part.

The Hadoop Image Processing Framework is intended to be extremely
simple to use. The framework strictly adheres to Java file writing and
reading techniques, extending those conventions to the Hadoop
framework's notion of reading and writing bundle files.  In this way,
working with Hadoop image processing is exactly like working on a
single system.  The main aim of the framework is to provide useful
software abstractions such that programming on a Hadoop cluster is
equivalent to programming on a single computer.  Listing 1
demonstrates the similarity between Java code written for a single
machine and the same code using the image processing framework.  The
framework's use of ordinary Java conventions helps in understanding
the software engineering process and reduces complexity.  Listing 2
demonstrates the mechanism for setting the image headers of processed
images.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\textwidth]{files-chart2}
  \caption{Comparing file writers and readers with and without using
    framework}
  \label{fig:files-chart}
\end{figure}

\begin{lstlisting}[caption = Comparing File Writer instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]

//Sample File Writing in Java

File file = new File("abc.txt");
FileWriter fw=new FileWriter(file);
fw.append(val)
fw.close();

//Sample BundleFile Writing in Hadoop using Framework

BundleFile file = new BundleFile("abc.seq");
SequenceBundleWriter sbw=new SequenceBundleWriter(file);
sbw.append(himage);
sbw.close();

\end{lstlisting}

\begin{lstlisting}[caption = Setting image headers for processed images using Hadoop Image Processing Framework]

//input - HImage is sent as input

CannyEdge canny = new CannyEdge(input);
canny.process();
HImage himage = canny.getProcessedImage();
himage.setImageHeader(input.getImageHeader());

\end{lstlisting}
\section{Conclusion}
This paper has described our Hadoop Image Processing Framework for
implementing large scale image processing applications.  The framework
is designed to abstract the technical details of Hadoop's powerful
MapReduce system and provide an easy mechanism for users to process
large image datasets. We provide software machinery for storing images
in the various Hadoop file formats and efficiently accessing the Map
Reduce pipeline.  By providing interoperability between different
image data types we allow the user to leverage many different
open-source image processing libraries. Finally, we provide the means
to preserve image headers throughout image manipulation process,
retaining useful and valuable information for image processing and
vision applications.

With these features, the framework provides a new level of
transparency and simplicity for creating large-scale image processing
applications on top of Hadoop's MapReduce framework.  We demonstrate
the power and effectiveness of the framework in terms of performance
enhancement and complexity reduction.  The Hadoop Image Processing
Framework should greatly expand the population of software developers
and researchers easily able to create large-scale image processing
applications.

\section{Future Work}
In the near future we hope to extend the framework into a full-fledged
multimedia processing framework. We would like to improve the
framework to handle audio and video processing over Hadoop with
similar ease. We also intend to add a CUDA module to allow processing
tasks to make use of machines' graphics cards. Finally, we intend to
develop our system into a highly parallelized open-source Hadoop
multimedia processing framework, providing web-based graphical user
interfaces for image processing applications.

\bibliography{references}
\bibliographystyle{IEEEtran}

\end{document}